Supervised Reinforcement Learning with Partial States: \\ a Teaching Method for Action Selection for HRI

### 2017-07-21 paper.tex
Pass on Partial States

\section{Partial-State Supervised Reinforcement Learning}
\begin{comment}
Humans evolve in a highly complex environment with a nearly infinite number of
dimensions originating from different modality through our senses. But, despite
this complexity, we can select actions according to important features of the
state, regardless of other parts of it.

Taking inspiration from this observation, we argue that focusing on partial states, rather
than full states can both increase the speed of learning for an agent and
provide some of the generalisation required to interact in complex environment.
Guidance from a human can highlight important features in the state space and
associate them to actions. This shifts the state action pair paradigm to a
partial-state action pair, and in the case of RL, the tuple: partial-state,
action and reward. 
\end{comment}

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. 
A classical approach is to use a first layer of neural
network or to use deep reinforcement learning which relies on deep neural
networks to learn a mapping state action Qvalue for example. 
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, as
discussed previously, in many application, these amounts of data are not
possible to be obtained and methods to generalise quickly are more desirable.

@@

In this paper we propose to use a human to highlight the relevant features of
the environment to reduce the state dimension of the points stored in memory.
We introduce the concept of partial-state, a sliced version of the state defined
only on a subset of the dimensions of the state. This shifts the (state,action)
pair paradigm to (partial-state,action)  
and in the case of RL, the tuple (state, action, reward) to (partial-state,
action, reward).

@@

In addition to these actions, the supervisor can use the GUI to select features
of the environment that can be sent to the algorithm to reduce the state associated
to this action to a partial-state. Similarly, the algorithm can indicate which
parts of the states have been used for the selection to the supervisor who can
correct them while the action is being executed.

@@

An expert supervises the agent actions, and can assign rewards to actions and
highlight the part of the states which are relevant to assigning this reward to
this partial-state.

@@

For this paper, we will reuse the formalism of rewardless Markov Decision 
Process to identify the different elements of our system. The agent has access
to a set of actions $\mathcal{A}$, a state $\mathcal{S} \in [0,1]^{n}$. We
also define the partial-states $\mathcal{s'} \in [0,1]^{n'}$ with $n' \leq n$ as a
slice of $\mathcal{S}$, a subset of $\mathcal{S}$ where some dimensions have
been removed. 
%With each dimension of S representing features in the environment.

At the instant $t$, the agent has access the state $s$, an action $a$, evaluated
with a reward $r$ associated to the partial-state $s'$. For example, a state $s$ could be
defined in 4 dimensions such as ${s=[1,0.2,0,0.5]}$, and $s'$ in two dimensions with
${s'=[-,0.2,0,-]}$ with symbol '$-$' reprensenting the dimensions removed. The
action $a$ has been evaluated $r$ in the partial-state $s'$.

To each action
$a \in \mathcal{A}$ we can associate a set $\mathcal{C}_{a}$ of pairs $(s',r)$
representing the rating done by the supervisor to action $a$ with feature
highlighted for $s'$. When adding a new pairt $(s',r)$, we can discard all the
previous pairs with an identical $s'$ to represent the evolution of the policy
evaluation by the supervisor.

\begin{algorithm}
    \DontPrintSemicolon
    \SetKwInOut{Input}{inputs}\SetKwInOut{Output}{output}
    \Input{Current state $s$, set of  $(a,s',r)$}
    \Output{selected action $\pi(s)$}
    \ForEach{$a \in \mathcal{A}$}{
        \ForEach{$(s'-r) \in \mathcal{C}_{a}$}{
            compute similarity $\Delta$ between $s$ and $s'$:
            $\Delta(s')=1-\frac{\sum_{i}^{n'}(s'(i)-s(i))^{2}}{n'}$
        }
        compute expected reward $\hat{r}(a)$:
        $\hat{r}(a) = max(\Delta) \cdot r(arg\, max_{s'} \Delta(s'))$\\
        with $r(\Delta(s'))$ reward of $(s'-r)$ pair
    }
    $\pi(s) = arg\, max_{a} \hat{r}(a)$

    \caption{Algorithm for selecting an action based on previous
    partial-state action rewards tuples and current state}
    \label{algo}
\end{algorithm}

When facing a new state $s$ where an action has to be
selected, the agent can select an action following Algorithm~\ref{algo}. For
each action $a \in \mathcal{A}$, we take the pair $(s',r)$ with the closest $s'$
to the current state (as defined by normalised distance over the dimensions
where $s'$ is defined). That way, each action $a$ can be associated to an
expected reward defined by the product between the similarity of the closest
partial-state known for $a$ and the reward obtained in that partial-state.
Finally, the action with the highest expected reward can be selected.

When selecting an action with following Algorithm~\ref{algo}, the agent can also use
the dimensions of the similar state $s'$ used for the selection to indicate the
supervisor which parts of the state have been used for the action selection.

@@

SPARC has been showed to be compatible with RL in \cite{senft2017supervised},
and can also be easily combined with the approach presented in paper using
partial states. For example, when selecting an action for the robot to be
executed, the supervisor can also select which features in the environment
should be selected in the partial state, and this associate the reward +1
to this action in this partial state. Similarly when an action is
proposed
to the supervisor the dimensions features represented by the dimensions of the
closest partial state can be explosed to the supervisor as a way to explain why
this action has been selected. Facing this, the supervisor can not react,
letting the action being executed and associating a reward of +1 to the partial
state and the action proposed, the supervisor could also change the partial
state to correct the features related to this action or cancel it, preventing the
execution and associating a reward of -1.

>>>>

\section{Partial State Supervised Reinforcement Learning}
\begin{comment}
Humans evolve in a highly complex environment with a nearly infinite number of
dimensions originating from different modality through our senses. But, despite
this complexity, we can select actions according to important features of the
state, regardless of other parts of it.

Taking inspiration from this observation, we argue that focusing on partial states, rather
than full states can both increase the speed of learning for an agent and
provide some of the generalisation required to interact in complex environment.
Guidance from a human can highlight important features in the state space and
associate them to actions. This shifts the state action pair paradigm to a
partial state action pair, and in the case of RL, the tuple: partial state,
action and reward. 
\end{comment}

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. 
A classical approach is to use a first layer of neural
network or to use deep reinforcement learning relying on deep neural
networks to learn a mapping state action Qvalue for example. 
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, 
 in many application, especially HRI, these amounts of data are not
obtainable due to time or unconsistency and methods to generalise with a lower
number of inputs are required.

@@

In this paper we propose to use a human to highlight the relevant features of
the environment to reduce the state dimension of the points stored in memory.
We introduce the concept of partial state, a sliced version of the state defined
only on a subset of the dimensions of the state. This shifts the (state,action)
pair paradigm to (partial state,action)  
and in the case of RL, the tuple (state, action, reward) to (partial state,
action, reward) and as such allows to compare the current state and the
datapoints only on relevant features for action selection.

@@

In addition to these actions, the supervisor can use the GUI to select features
of the environment that can be sent to the algorithm to reduce the state associated
to this action to a partial state. Similarly, the algorithm can indicate which
parts of the states have been used for the selection to the supervisor who can
correct them while the action is being executed.

@@

An expert supervises the agent actions, and can assign rewards to actions and
highlight the parts of the states which are relevant to assigning this reward to
this partial state.

@@

For this paper, we will reuse the formalism of rewardless Markov Decision 
Process to identify the different elements of our system. The agent has access
to a set of actions $\mathcal{A}$, a state $\mathcal{S} \in [0,1]^{n}$. We
also define the partial states $\mathcal{S'} \in [0,1]^{n'}$ with $n' \leq n$ as a
slice of $\mathcal{S}$, a subset of $\mathcal{S}$ where some dimensions have
been removed. 
%With each dimension of S representing features in the environment.

At the instant $t$, the agent has access the state $s$, an action $a$, evaluated
with a reward $r$ associated to the partial state $s'$. For example, a state $s$ could be
defined in 4 dimensions such as ${s=[1,0.2,0,0.5]}$, and $s'$ in two dimensions with
${s'=[-,0.2,0,-]}$ with symbol '$-$' reprensenting the dimensions removed. The
action $a$ has been evaluated $r$ in the partial state $s'$.

To each action
$a \in \mathcal{A}$ we can associate a set $\mathcal{C}_{a}$ of pairs $(s',r)$
representing the rating done by the supervisor to action $a$ with feature
highlighted for $s'$. When adding a new pairt $(s',r)$, we can discard potential
previous pairs with an identical $s'$ to represent the evolution of the policy
evaluation by the supervisor.

\begin{algorithm}
    \DontPrintSemicolon
    \SetKwInOut{Input}{inputs}\SetKwInOut{Output}{output}
    \Input{Current state $s$, set of  $(a,s',r)$}
    \Output{selected action $\pi(s)$}
    \ForEach{$a \in \mathcal{A}$}{
        \ForEach{$p=(s',r) \in \mathcal{C}_{a}$}{
            compute similarity $\Delta$ between $s$ and $s'$:
            $\Delta(p)=1-\frac{\sum_{i}^{n'}(s'(i)-s(i))^{2}}{n'}$
        }
        compute expected reward $\hat{r}(a)$ for taking $a$ in $s$:
        $\hat{r}(a) = max(\Delta) \cdot r(arg\, max_{p} \Delta(p))$\\
        with $r(\Delta(p))$ the reward of $p = (s',r)$ 
    }
    Select the action with the maximum expected reward:
    $\pi(s) = arg\, max_{a} \hat{r}(a)$

    \caption{Algorithm for selecting an action based on previous
    partial state action rewards tuples and current state}
    \label{algo}
\end{algorithm}

When facing a new state $s$ where an action has to be
selected, the agent can select an action following Algorithm~\ref{algo}. For
each action $a \in \mathcal{A}$, we take the pair $(s',r)$ with the closest $s'$
to the current state (as defined by normalised distance over the dimensions
where $s'$ is defined). That way, each action $a$ can be associated to an
expected reward defined by the product between the similarity of the closest
partial state known for $a$ and the reward obtained for executing $a$ in that partial state.
Finally, the action with the highest expected reward can be selected.

When selecting an action with following Algorithm~\ref{algo}, the agent can also use
the dimensions of the close state $s'$ used for the selection to indicate the
supervisor which parts of the state have been used for the action selection.

@@

SPARC has been showed to be compatible with RL in \cite{senft2017supervised},
and can also be easily combined with the approach presented in this paper using
partial states. For example, when selecting an action for the robot to be
executed, the supervisor can also select which features in the environment
should be selected in the partial state, and this associates the reward +1
to this action in this partial state. Similarly when an action is
proposed to the supervisor the features represented by the dimensions of the
closest partial state can be exposed to the supervisor as a way to explain why
this action has been selected. Facing this, the supervisor can not react,
letting the action being executed and associating a reward of +1 to the partial
state and the action proposed, change the partial state to correct the features
related to this action or cancel it, preventing the execution and associating a
reward of -1.
